---
title: "Single Cell Analysis Report"
author: "scShinyHub"
date: "3/8/2018"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 7
    fig_width: 9
    number_sections: yes
    toc: yes
params:
#__PARAMPLACEHOLDER__
---

```{r checkDEBUG, include=FALSE}
if (!exists("DEBUG")) {
  DEBUG <- TRUE
  # DEBUG=FALSE
}
if (!exists("DEBUGSAVE")) {
  # DEBUGSAVE=TRUE
  DEBUGSAVE <- FALSE
}
```

```{r loadData, include=FALSE}
require(shiny)
require(shinyTree)
require(shinyBS)
require(plotly)
require(shinythemes)
require(ggplot2)
require(DT)
require(pheatmap)
require(threejs)
require(sm)
require(RColorBrewer)
require(mclust)
require(reshape)
# require(cellrangerRkit)
# require(SCORPIUS)
require(ggplot2)
require(knitr)
require(kableExtra)
require(shinyWidgets)
# require(scater)

# params only exsits if called from somewhere with parameters
if (exists("params") & is.list(params)) {
  cat(file = stderr(), paste("params:", params$calledFromShiny, "\n"))
  cat(file = stderr(), paste("params exists:", "calledFromShiny" %in% names(params), "\n"))
  LOCALEXECUTION <- FALSE
  if (DEBUGSAVE) {
    base::save(file = "~/scShinyHubDebug/tempReport-rmd.RData", list = c(ls()))
  }
} else {
  # rm(list = ls())
  source("serverFunctions.R")
  source("reactives.R", local = TRUE)
  uiFiles <- dir(path = "contributions", pattern = "reactives.R", full.names = TRUE, recursive = TRUE)
  for (fp in uiFiles) {
    if (DEBUG) cat(file = stderr(), paste("loading: ", fp, "\n"))
    source(fp, local = TRUE)
  }
  cp = load("~/scShinyHubDebug/tempReport.RData")
  params <- myparams
  LOCALEXECUTION <- TRUE # to know that we are debugging.
}
```


__LOAD_REACTIVES__

# parameters

```{r setup, include=TRUE, echo=FALSE, warning=FALSE}
input <- params
print(str(input$file1$name))

if (file.exists(input$file1$datapath)) {
  load(input$file1$datapath)
} else {
  cat(file = stderr(), paste("Error input file not existing\n"))
}
# print(ls())
# will be printed on console, for debugging purposes only
cat(file = stderr(), paste("env:", ls(), "\n"))
cat(file = stderr(), paste("input:", names(input), input, "\n"))
# cat(file = stderr(), paste("dataTables:", names(dataTables),"\n"))
```



```{r check Variables, include=FALSE}
# TODO this needs to go to individual packages

if (is.null(input$cluster)) {
  input$cluster <- "0"
  print("Warning: setting cluster to 0")
}
if (is.null(input$cluster5)) {
  input$cluster5 <- "0"
  print("Warning: setting cluster5 to 0")
}
if (is.null(input$clusters)) {
  input$clusters <- "0"
  print("Warning: setting clusters to 0")
}
if (is.null(input$sCA_dgeClustersSelection)) {
  input$sCA_dgeClustersSelection <- "0"
  print("Warning: setting sCA_dgeClustersSelection to All")
}
if (is.null(input$clusters2)) {
  input$clusters2 <- "0"
  print("Warning: setting clusters2 to 0")
}
if (is.null(input$clusters3)) {
  input$clusters3 <- "0"
  print("Warning: setting clusters4 to 0")
}
if (is.null(input$DE_clusterSelectionPanelPlot)) {
  input$DE_clusterSelectionPanelPlot <- "All"
  print("Warning: setting clusters4 to All")
}
```


```{r inputData, echo=FALSE}
# if ( !exists("dataTables")) {# needed when working with ~/scShinyHubDebug/reports.Data
cat(file = stderr(), paste("loading dataTables\n"))
# in case the file is not in the root directory is has to be manually loaded during debugging
# input$file1 = data.frame("datapath" = dir(path = "~/Rstudio/scShinyHubData/", pattern="anna1805.sp.Rds", full.names = T), stringsAsFactors = F)
# inFile = input$file1
dataTables <- inputDataFunc(input$file1)
# }
```

```{r useCells, echo=FALSE}
geneNames <- input$minExpGenes
rmCells <- input$cellsFiltersOut
rmPattern <- input$cellPatternRM
keepCells <- input$cellKeep
# cat(file = stderr(), paste("dataTables:", names(dataTables),"\n"))
useCells <- useCellsFunc(dataTables,
  geneNames = input$minExpGenes,
  rmCells = input$cellsFiltersOut,
  rmPattern = input$cellPatternRM,
  keepCells = input$cellKeep,
  cellKeepOnly = input$cellKeepOnly
)
```

```{r useGenes, echo=FALSE}
# base::save(file = "~/scShinyHubDebug/report.useGenes.RData", list = ls())
ipIDs <- input$selectIds
genesKeep <- input$genesKeep
geneListSelection <- input$geneListSelection

useGenes <- useGenesFunc(dataTables, ipIDs, geneListSelection, genesKeep, geneLists)
```


```{r featureDataReact, echo=FALSE}

featureDataReact <- dataTables$featuredata[useGenes, ]
featureData <- dataTables$featuredata[useGenes, ]
```

```{r scEx, echo=FALSE}
if (DEBUG) cat(file = stderr(), "scEx\n")
minGene <- input$minGenesGS # min number of reads per gene
minG <- input$minGenes # min number of reads per cell
maxG <- input$maxGenes # max number of reads per cell

scEx <- scExFunc(scExOrg = dataTables$scEx, useCells = useCells, useGenes = useGenes, minGene = minGene, minG = minG, maxG = maxG)

if (DEBUG) cat(file = stderr(), "scEx:DONE\n")
if (DEBUG) cat(file = stderr(), paste("scEx:", class(scEx), "\n"))
```


```{r medianENSG, echo=FALSE}
medianENSG <- medianENSGfunc(as.matrix(assays(scEx)[["counts"]]))
```


```{r medianUMI, echo=FALSE}
medianUMI <- medianUMIfunc(as.matrix(assays(scEx)[["counts"]]))
```

```{r groupNames, echo=FALSE}
groupNames <- list()
groupNames$namesDF <- gNames
```


```{r scEx_log, eval=FALSE, include=FALSE}
# scEx_log is the normalized version of the data. This stored in a separate RData
# and thus doen't need to be recomputed.
if (DEBUG) cat(file = stderr(), "scEx_log\n")
# use_genes <- get_nonzero_genes(scEx)
# scEx_bcnorm <- normalize_barcode_sums_to_median(scEx)
# scEx_log <- log_gene_bc_matrix(scEx_bcnorm,base=10)


normMethod <- input$normalizationRadioButton
# can only be done in reactive context...
# thus, here the minimal version for debugging
    # use_genes <- get_nonzero_genes(scEx)
    # scEx_bcnorm <- normalize_barcode_sums_to_median(scEx)
    # scEx_log <- log_gene_bc_matrix(scEx_bcnorm, base = 10)

scEx_log <- do.call(normMethod, args = list())
if (DEBUG) cat(file = stderr(), "scEx_log:Done\n")
```


```{r scEx_matrix, include=FALSE}

scEx_matrix <- as.matrix(assays(scEx)[["counts"]])
```


```{r rawNormalization, include=FALSE}
rawNormalization <- scEx
```

```{r scExLogMatrixDisplay, include=FALSE}
retVal <- as.data.frame(as.matrix(assays(scEx_log)[[1]]))
rownames(retVal) <- make.names(rowData(scEx_log)$symbol, unique = TRUE)

scExLogMatrixDisplay <- retVal
```


```{r pca, eval=TRUE, echo=FALSE}
if (DEBUG) cat(file = stderr(), "pca\n")

pca <- pcaFunc(scEx_log)
```
Creating  `r input$kNr` clusters.

```{r kmClustering, eval=TRUE, echo=FALSE}
if (DEBUG) cat(file = stderr(), "kmClustering\n")
seed <- 1
  # retVal <- kmClusteringFunc(pca, seed = seed, kNr = kNr)
  retVal <- scranCluster(
    pca, scEx_log, seed, clusterSource,
    geneSelectionClustering, minClusterSize, clusterMethod, 
    featureData
  )

if (DEBUG) cat(file = stderr(), "kmClustering:done\n")
```

Creating `r input$gQC_tsneDim` tSNE clusters.

```{r tsne, eval=TRUE, echo=FALSE}
if (DEBUG) cat(file = stderr(), "tsne\n")
gQC_tsneDim <- input$gQC_tsneDim
gQC_tsnePerplexity <- input$gQC_tsnePerplexity
gQC_tsneTheta <- input$gQC_tsneTheta
gQC_tsneSeed <- input$gQC_tsneSeed
set.seed(seed = gQC_tsneSeed)
tsne <- run_tsne(pca,
  dims = gQC_tsneDim,
  perplexity = gQC_tsnePerplexity,
  theta = gQC_tsneTheta, check_duplicates = FALSE
)

if (DEBUG) cat(file = stderr(), "tsne: done\n")
```


```{r projections, eval=FALSE, include=FALSE}
# TODO not working with current projection model and debugging.
# the correct projections are used via saving/loading of variables during report creation in server.R
#     base::save(file = tmpPrjFile, list = c("projections", "scEx_log", "gNames"))
#     reactiveFiles <- paste0(reactiveFiles, "load(file=\"", tmpPrjFile, "\")\n", collapse = "\n")

if (DEBUG) cat(file = stderr(), "projections\n")
clustering <- kmClustering
projections <- data.frame(tsne$Y)
colnames(projections) <- paste0("tsne", c(1:ncol(projections)))
# projections$dbCluster <- factor(clustering$kmeans_10_clusters$Cluster - 1)
projections$dbCluster <- factor(clustering$kmeans_10_clusters$Cluster )

projections <- cbind(projections, data.frame(pca$x[, c(1, 2, 3)]))

rownames(projections) <- clustering$kmeans_10_clusters$Barcode
samp <- gsub(".*-(.*)", "\\1", rownames(projections))
if (length(levels(as.factor(samp))) > 1) {
  projections$sample <- samp
} else {
  projections$sample <- "1"
}

n <- length(projectionFunctions)
for (proj in projectionFunctions) {
  if (DEBUG) cat(file = stderr(), paste("forceCalc ", proj[1], "\n"))
  assign("tmp", eval(parse(text = paste0(proj[2], "()"))))
  cn <- make.names(c(colnames(projections), make.names(proj[1])))
  if (length(tmp) == 0) {
    next()
  }
  if (ncol(projections) == 0) {
    projections <- data.frame(tmp = tmp)
  } else {
    if (nrow(projections) == length(tmp)) {
      projections <- cbind(projections, tmp)
    }
  }
  colnames(projections) <- cn
}
projections$UmiCountPerGenes <- 0
projections$UmiCountPerGenes2 <- 0

if (DEBUG) cat(file = stderr(), "projections: done\n")
```

```{r sampleInfo, include=FALSE}
sampleInfo <- sampleInfoFunc(scEx)
```

```{r inputSample, include=FALSE}
sampInf <- gsub(".*-(.*)", "\\1", dataTables$scEx$barcode)
inputSample <- data.frame(
  cellName = colnames(dataTables$scEx),
  sample = sampInf,
  ncells = colSums(as.matrix(assays(dataTables$scEx)[[1]]))
)
```


```{r log2cpm, eval=TRUE, include=FALSE}
if (DEBUG) cat(file = stderr(), "log2cpm\n")
log2cpm <- log2cpm <- as.data.frame(as.matrix(assays(scEx_log)[[1]]))
if (DEBUG) cat(file = stderr(), "log2cpm:Done\n")
```


```{r summaryStats, eval=TRUE, include=TRUE}


line1 <- paste("No. of cells:", dim(log2cpm)[2], sep = "\t")
line2 <- paste("Median UMIs:", medianUMI, sep = "\t")
line3 <- paste("Median Genes:", medianENSG, sep = "\t")
line5 <- paste("No. of reads:", dim(log2cpm)[1], sep = "\t")
HTML(
  paste0(
    "Summary statistics of this dataset:", "<br/>", "<br/>",
    line1, "<br/>",
    line2, "<br/>",
    line3, "<br/>",
    line5
  )
)
```

# Input/General

input file: **`r input$file1$name`**

Description:

**`r input$descriptionOfWork`**

## Normalization used

Normalization method used: **`r input$normalizationRadioButton`**.

## Variables used

Variables that are used can be found here: **[variables.used.txt](variables.used.txt)


__CHILDREPORTS__

```{r variables, include=FALSE, results='asis'}
# cat(file = stderr(), paste(ls(), collapse = "\n"))
# save(file = "~/scShinyHubDebug/report.vars.RData", list = ls())
# load(file = "~/scShinyHubDebug/report.vars.RData")
fileConn <- (paste0(input$reportTempDir, "/variables.used.txt"))
printVars <- function(x, fileConn) {
  for (varN in names(x)) {
    # if (is.data.frame(x[[varN]])) {
    cat(file = fileConn, x = paste(varN, "=", paste(x[[varN]], collapse = "\n"), collapse = "\n"), append = TRUE)
    cat(file = fileConn, x = "\n\t", append = TRUE)
  }
}
printVars(params, fileConn)
# close(fileConn)
```

# Session information

```{r sessionInfo, echo=FALSE}
if (!LOCALEXECUTION) {
  # save(file = "~/scShinyHubDebug/report.RData", list=ls())
}
sessionInfo()
```


