---
title: "coExpression"
author: "Bernd Jagla"
date: "3/23/2018"
output: html_document
---

# Co-expression

```{r bj-coExp-setup, echo=FALSE, warning=FALSE}
if (exists("params")) {
  cat(file = stderr(), paste("Scater Plot report\n"))
} else {
  rm(list = ls())
  load("~/scShinyHubDebug/report.RData")
  cat(file = stderr(), getwd())
  
  require(shiny)
  require(plotly)
  require(shinythemes)
  require(ggplot2)
  require(DT)
  require(pheatmap)
  require(threejs)
  require(sm)
  require(RColorBrewer)
  require(mclust)
  require(reshape)
  # require(cellrangerRkit)
  # require(SCORPIUS)
  require(ggplot2)
  require(knitr)
  require(kableExtra)
  require(shinyWidgets)
  require(scater)
  
  source("../../../serverFunctions.R")
  source("../../../reactives.R", local = TRUE)
  # source("reactives.R", local = TRUE)
  
  LOCALEXECUTION <- TRUE # to know that we are debugging.
  useCells <- useCellsFunc(dataTables,
                           geneNames = input$minExpGenes,
                           rmCells = input$cellsFiltersOut,
                           rmPattern = input$cellPatternRM,
                           keepCells = input$cellKeep,
                           cellKeepOnly = input$cellKeepOnly
  )
  ipIDs <- input$selectIds
  geneListSelection <- input$geneListSelection
  minGene <- input$minGenesGS
  
  useGenes <- useGenesFunc(dataTables, useCells, ipIDs, geneListSelection, minGene)
  featureDataReact <- dataTables$featuredata[useGenes, ]
  featureData <- dataTables$featuredata[useGenes, ]
  scEx <- dataTables$scEx[useGenes, useCells]
  scEx_log <- dataTables$scEx_log[useGenes, useCells]
  log2cpm <- dataTables$log2cpm[useGenes, useCells]
}
```

## All clusters 

### Heatmap

```{r coE_heatmapFunc, echo=FALSE}
require(pheatmap)

scEx_matrix <- assays(scEx_log)[[1]]
genesin <- input$heatmap_geneids
sampCol = sampleCols$colPal

genesin2 <- geneName2Index(genesin, featureData)
genesin = (paste0(featureData[genesin2[which (genesin2 %in% rownames(scEx_matrix) )],"Associated.Gene.Name"], collapse =",", sep=""))

heatmapData <- coE_heatmapFunc(
  featureData = featureData, scEx_matrix = scEx_matrix,
  projections = projections, genesin = genesin, cells = colnames(scEx_matrix),
  sampCol = sampCol
)

addColNames <- input$"coExpHeatmapModule-ColNames"
orderColNames <- input$"coExpHeatmapModule-orderNames"
moreOptions <- input$"coExpHeatmapModule-moreOptions"
proje <- projections

heatmapData$filename <- NA
if (is.null(moreOptions)) moreOptions <- FALSE

if (length(addColNames) > 0 & moreOptions) {
  heatmapData$annotation_col <- proje[rownames(heatmapData$annotation_col), addColNames, drop = FALSE]
}
if (sum(orderColNames %in% colnames(proje)) > 0 & moreOptions) {
  heatmapData$cluster_cols <- FALSE
  heatmapData$mat <- heatmapData$mat[, rownames(dfOrder(proje, orderColNames)), drop = FALSE]
}

do.call(pheatmap, heatmapData)
```

## Selected

### 2D plot with selection of cells 

geneNames <- `r input$"selected-geneIds"`

clId <- `r input$"selected-clusters"`

dimX <- `r input$"selected-dimension_x"`

dimY <- `r input$"selected-dimension_y"`



```{r selectd2D, echo=FALSE}
g_id <- input$gene_id
geneNames <- input$"selected-geneIds"
#TODO geneNames2 is just added to have a value, need to check how this should behave.
geneNames2 <- input$"selected-geneIds"
clId <- input$"selected-clusters"
dimX <- input$"selected-dimension_x"
dimY <- input$"selected-dimension_y"
legend.position <- "none"
grpN <- make.names(input$groupName)
cat(file = stderr(), paste("2Dprojection", c(g_id, geneNames, dimX, dimY, clId, grpN, legend.position)))

# save(file = "~/scShinyHubDebug/bj.DE.2Dprojection.Rdata", list = ls())
# load(file = "~/scShinyHubDebug/bj.DE.2Dprojection.Rdata")

plot2Dprojection(scEx_log, scEx, projections, g_id, featureData, geneNames, geneNames2, dimX, dimY, clId, grpN, legend.position, 
                 grpNs = groupNames$namesDF)
```


#### Heatmap selected

callModule(
pHeatMapModule, 
"heatmapSelectedModule", 
heatmapSelectedReactive
)

heatmap_geneids2 `r input$heatmap_geneids2`

'selected-geneIds' `r input$'selected-geneIds'`

'selected-clusters' `r input$'selected-clusters'`

'heatmapSelectedModule-ColNames' `r input$'heatmapSelectedModule-ColNames'`

orderColNames : `r input$'heatmapSelectedModule-orderNames'`

moreOptions <- `r input$'heatmapSelectedModule-moreOptions'`


```{r heatmapSelectedModule, echo=FALSE, eval=TRUE}
# save(file = "~/scShinyHubDebug/selectedHeatmap.1.RData", list = c(ls(), ls(envir = globalenv())))
genesin <- input$heatmap_geneids2

# get selected cells
brushedPs <- input$"selected-b1"
geneNames <- input$'selected-geneIds'
inpClusters <- input$"selected-clusters"
dimX <- brushedPs$mapping$x
dimY <- brushedPs$mapping$y
sampCol = sampleCols$colPal

if(!is.null(brushedPs)){
  
  # geneid <- geneName2Index(geneNames, featureData)
  projections <- updateProjectionsWithUmiCount(dimX = dimX, dimY=dimY, 
                                               geneNames = geneNames,
                                               featureData = featureData,
                                               scEx = scEx, projections = projections)
  
  subsetData <- subset(projections, dbCluster %in% inpClusters)
  cells.names <- shiny::brushedPoints(subsetData, brushedPs)
  
  cells.1 <- rownames(cells.names)
  heatmapData <- coE_heatmapFunc(featureData,
                                 scEx_matrix,
                                 projections,
                                 genesin,
                                 cells = cells.1,
                                 sampCol = sampCol
  )
  
  # print heatmap
  addColNames <- input$"heatmapSelectedModule-ColNames"
  orderColNames <- input$"heatmapSelectedModule-orderNames"
  moreOptions <- input$"heatmapSelectedModule-moreOptions"
  proje <- projections
  if (DEBUGSAVE) {
    save(file = "~/scShinyHubDebug/selectedHeatmap.RData", list = c(ls(), ls(envir = globalenv())))
  }
  # save(file = "~/scShinyHubDebug/selectedHeatmap.3.RData", list = c(ls(), ls(envir = globalenv())))
  # load(file = "~/scShinyHubDebug/selectedHeatmap.3.RData")
  heatmapData$filename <- NA
  if (is.null(moreOptions)) moreOptions <- FALSE
  
  if (length(addColNames) > 0 & moreOptions) {
    heatmapData$annotation_col <- proje[rownames(heatmapData$annotation_col), addColNames, drop = FALSE]
  }
  if (sum(orderColNames %in% colnames(proje)) > 0 & moreOptions) {
    heatmapData$cluster_cols <- FALSE
    heatmapData$mat <- heatmapData$mat[, rownames(dfOrder(proje, orderColNames)), drop = FALSE]
  }
  # save(file = "~/scShinyHubDebug/selectedHeatmap.4.RData", list = c(ls(), ls(envir = globalenv())))
  
  do.call(pheatmap, heatmapData)
}
```

#### table of selected cells

```{r topexpCells, echo=FALSE}
if (DEBUG) cat(file = stderr(), "output$topExpGenes\n")
coEtgPerc <- input$coEtgPerc
coEtgminExpr <- input$coEtgMinExpr
outMat <- NULL
if(!is.null(brushedPs)){
  
  scCells <- cells.1
  
  if (DEBUGSAVE) {
    save(file = "~/scShinyHubDebug/output_topExpGenes.RData", list = c("scCells", ls()))
  }
  # load(file="~/scShinyHubDebug/output_topExpGenes.RData")
  # we only work on genes that have been selected
  mat <- assays(scEx_log)[[1]][, scCells]
  # only genes that express at least coEtgminExpr UMIs
  mat[mat < coEtgminExpr] <- 0
  # only genes that are expressed in coEtgPerc or more cells
  allexpressed <- Matrix::rowSums(mat > 0) / length(scCells) * 100 >= coEtgPerc
  mat <- mat[allexpressed, ]
  
  cv <- function(x) {
    sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
  }
  matCV <- apply(mat, 1, cv)
  # top.genes <- as.data.frame(exprs(scEx_log))
  maxRows <- min(nrow(mat), 200)
  top.genesOrder <- order(matCV, decreasing = TRUE)[1:maxRows]
  # 2do dim is not correct produces problmes
  
  if (dim(mat)[1] > 0) {
    mat <- mat[top.genesOrder, ]
    fd <- featureData[rownames(mat), c("Associated.Gene.Name", "Description")]
    matCV <- matCV[rownames(mat)]
    fd <- cbind(fd, matCV)
    colnames(fd) <- c("gene", "description", "CV")
    outMat <- cbind(fd, as.matrix(mat))
    rownames(outMat) <- make.names(as.character(outMat$gene), unique = TRUE)
    DT::datatable(outMat,
                  rownames = F, filter = "top",
                  options = list(
                    orderClasses = TRUE,
                    autoWidth = TRUE
                  )
    )
  }
}
```

```{r saveTOPGENES, echo=FALSE}
# save full table in tmp folder to be included in report
write.csv(outMat, file = paste0(input$reportTempDir, "/output_topExpGenes.csv"))

```

Full table can be found here: [output_topExpGenes.csv](output_topExpGenes.csv)


### binarized

genesin <- `r input$mclustids`

cl3 <- `r input$clusters3`

dimx3 <- `r input$dimension_x3`

dimy3 <- `r input$dimension_y3`

not really working on this since I don't know if we really need it what this is about..

```{r coEx-binarized, eval=FALSE, echo=FALSE}

genesin <- input$mclustids
cl3 <- input$clusters3
dimx3 <- input$dimension_x3
dimy3 <- input$dimension_y3
# posCells <- positiveCells$positiveCells # we use this variable to be able to save the global variable in this context
# posCellsAll <- positiveCells$positiveCellsAll


if (DEBUGSAVE) {
  save(file = "~/scShinyHubDebug/plotCoExpression.RData", list = c(ls(), ls(envir = globalenv())))
}
# load(file="~/scShinyHubDebug/plotCoExpression.RData")
p1 <- plotCoExpressionFunc(
  featureData,
  scEx_log,
  upI,
  projections,
  genesin,
  cl3,
  dimx3,
  dimy3
)
return(p1)
```

## Violin plot coexpressed genes

geneListStr <- `r input$geneGrpVioIds`

projectionVar <- `r input$dimension_xVioiGrp`

minExpr <- `r input$coEminExpr`

```{r violinPlot, echo=FALSE}
geneListStr <- input$geneGrpVioIds
projectionVar <- input$dimension_xVioiGrp
minExpr <- input$coEminExpr

retVal <- geneGrp_vioFunc(
  genesin = geneListStr,
  projections = projections,
  scEx = scEx,
  featureData = featureData,
  minExpr = minExpr,
  dbCluster = projectionVar,
  showPermutations = input$showPermutations
)
retVal
```



